---
title: "Individual-subject example from the tutorial paper"
author: "Jeffrey Annis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
     %\VignetteIndexEntry{Examples from the tutorial paper}
     %\VignetteEngine{knitr::rmarkdown}
     %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
     collapse = TRUE,
     comment = "#>"
)
```
This vignette describes the individual-subject examples in the paper "Thermodynamic integration and Steppingstone sampling methods for computing Bayes factors: A tutorial" by Annis, Evans, Miller, and Palmeri (submitted). The paper comes bundled with the `powder` package. To view it:
```{r,eval=FALSE}
vignette("powder-paper",package = 'powder')
```
For the individual-subject examples, we used a dataset containing a single simulated subject in two conditions with 600 trials per condition. The dataset is the same as the "Simple" dataset used in Evans and Brown (2017). We fit two models: a "simple" model with no parameters varying across conditions, and a "complex" model with 3 different parameters varying across conditions: non-decision time `t0`, threshold `b`, and correct drift rate `vc`. Let's load the dataset and create the models.
```{r}
library(powder)
data('individual')
simple.individual.model = LBA.Individual$new(contaminant = list(pct=2,upper.bound=5))
complex.individual.model = LBA.Individual$new(b=TRUE,vc=TRUE,t0=TRUE,contaminant = list(pct=2,upper.bound=5))
```
Now that we have defined the models, let's use `powder` to sample from the power posteriors. We will sample from posteriors raised to 20 temperatures, each with 200 samples. Note, the `meltin` parameter describes how long to adapt to each temperature, akin to burnin.
```{r,warning=FALSE,eval=FALSE}
simple.individual.out = powder(model=simple.individual.model,data=individual,
                               num.temps=30,burnin=500,meltin=50,n.samples=100)
complex.individual.out = powder(model=complex.individual.model,data=individual,
                                num.temps=30,burnin=500,meltin=50,n.samples=100)
```
Now that we have the samples, let's take a look at the marginal likelihoods. Note, the harmonic mean and Steppingstone estimators are different from the other estimates. The harmonic mean tends to overestimate the marginal likelihood and the Steppingstone estimator is unstable because it is estimating p(D) rather than ln(p(D)) like the other estimators. It's advised to not use harmonic mean and to use the Log Steppingstone estimator instead of the Steppingstone estimator.
```{r,eval=FALSE}
simple.individual.ml = summary(simple.individual.out)
print(simple.individual.ml)

#                   Method         Value
# 1                     TI  5.619068e+02
# 2           TI Corrected  5.622413e+02
# 3          Harmonic Mean  5.756293e+02
# 4          Steppingstone 5.719795e+242
# 5      Log Steppingstone  5.621503e+02
# 6            TI Variance  2.447853e-03
# 7 Steppingstone Variance  1.628729e-02

complex.individual.ml = summary(complex.individual.out)
print(complex.individual.ml)

#                   Method         Value
# 1                     TI  5.549913e+02
# 2           TI Corrected  5.555068e+02
# 3          Harmonic Mean  5.765305e+02
# 4          Steppingstone 6.268520e+238
# 5      Log Steppingstone  5.556290e+02
# 6            TI Variance  3.266298e-03
# 7 Steppingstone Variance  2.450748e-02

```
Sampling from the power posteriors sequentially can be slow, especially for complex models. To sample from the posteriors in parallel, where each chain samples from a different target distribution use the argument `method='parallel`.
```{r,eval=FALSE}
simple.individual.out.par = powder(model=simple.individual.model,data=individual,
                               num.temps=30,burnin=500,n.samples=1000,method='parallel')
simple.individual.ml.par = marginal.likelihood(simple.individual.out.par)
print(simple.individual.ml.par)

#                   Method         Value
# 1                     TI  5.613116e+02
# 2           TI Corrected  5.616935e+02
# 3          Harmonic Mean  5.761761e+02
# 4          Steppingstone 2.700302e+242
# 5      Log Steppingstone  5.614802e+02
# 6            TI Variance  4.656865e-03
# 7 Steppingstone Variance  3.609845e-02

complex.individual.out.par = powder(model=complex.individual.model,data=individual,
                               num.temps=30,burnin=500,n.samples=2000,method='parallel')
complex.individual.ml.par = summary(complex.individual.out.par)
print(complex.individual.ml.par)

#                   Method         Value
# 1                     TI  5.531538e+02
# 2           TI Corrected  5.536072e+02
# 3          Harmonic Mean  5.765168e+02
# 4          Steppingstone 4.494607e+237
# 5      Log Steppingstone  5.531707e+02
# 6            TI Variance  4.026619e-03
# 7 Steppingstone Variance  7.006088e-02
```
It's easy to compute the Bayes factor in terms of the simple model over the complex model.
```{r,eval=FALSE}
bayes.factor(simple.individual.ml, complex.individual.ml)

#              Method Log.Bayes.factor
# 1                TI        6.9154542
# 2 Log Steppingstone        6.5213709
# 3     Harmonic Mean       -0.9012413

bayes.factor(simple.individual.ml.par, complex.individual.ml.par)

#              Method Log.Bayes.factor
# 1                TI        8.1578545
# 2 Log Steppingstone        8.3094503
# 3     Harmonic Mean       -0.3406929
```
